---
title: "GATS Methods"
format: html
editor: source
code-fold: true
---
```{r setup, message=F}
library(here)
library(tidyverse)
library(labelled)
library(kableExtra)
```

## General approach

The primary purpose of this analysis is to describe the prevalence of dual and poly-tobacco use in selected LMICs using the GATS surveys. 

Read in the GATS survey
```{r}
gd <- read_rds(here("data-clean", "gats-clean.rds")) %>%
  select(-hwarn) %>%
  remove_var_label() %>%
  drop_na() %>%
  mutate(poly = ifelse(numtob > 1, 1, 0))

kbl(head(gd)) %>% 
  kable_styling()
```

Although the overall size of the GATS survey dataset is large (N = ~230,000), the cell sizes for individual countries get quite small once we begin stratifying on demographic and other factors. For example

```{r}
gd %>% 
  count(country, male, agegp, wealth) %>% 
  pivot_wider(names_from = agegp, 
    values_from = n, names_prefix = "Age") %>% 
  pivot_wider(names_from = wealth, 
   values_from = starts_with("Age"), 
   names_prefix = "WQ") %>% 
  kbl() %>% kable_styling()
```

You can see, for example, when just stratifying by sex, wealth, and agegroup we have a number of cells <10 (Botswana) and others (Tanzania) that are <20. This leads to uncertainty and means we cannot, for example, generate estimates for important subpopulations within countries. 

## Statistical methods

### Multi-level Regression and Poststratification

One of the primary benefits of the MRP approach is that it allows for the estimation o

We can define our multi-level model as follows: 

$$
Pr(y_i = 1) = logit^{-1}(
\alpha_{\rm c[i]}^{\rm country}
+ \alpha_{\rm a[i]}^{\rm age}
+ \alpha_{\rm w[i]}^{\rm wealth}
+ \alpha_{\rm e[i]}^{\rm educ}
+ \beta^{\rm male} \cdot {\rm Male}_{\rm i} 
+ \alpha_{\rm g[i], r[i]}^{\rm male.eth}
+ \alpha_{\rm e[i], a[i]}^{\rm educ.age}
+ \alpha_{\rm e[i], r[i]}^{\rm educ.eth}
)
$$
Or this:

$$
\begin{aligned}
&\ \textbf{Probability of selection of alternative}_i \textbf{ in subject}_j \\
\text{Choice}_{i_j} \sim&\ \operatorname{Poisson}(\lambda_{i_j}) \\[10pt]
&\ \textbf{Model for probability of each option} \\
\log(\lambda_{i_j}) =&\ (\beta_0 + b_{0_j}) + \beta_1 \text{Dark}_{i_j} + \beta_2 \text{Soft}_{i_j} + \beta_3 \text{Nuts}_{i_j} \\[5pt]
b_{0_j} \sim&\ \mathcal{N}(0, \sigma_0) \qquad\qquad\quad \text{Subject-specific offsets from global choice probability} \\[10pt]
&\ \textbf{Priors} \\
\beta_0 \sim&\ \mathcal{N}(0, 3) \qquad\qquad\quad\ \ \text{Prior for global average choice probability} \\
\beta_1, \beta_2, \beta_3 \sim&\ \mathcal{N}(0, 3) \qquad\qquad\quad\ \ \text{Prior for candy feature levels} \\
\sigma_0 \sim&\ \operatorname{Exponential}(1) \qquad \text{Prior for between-subject variability}
\end{aligned}
$$
